\chapter{Conclusion}
\label{CHAP_SEVENTH}
\centerline{\rule{149mm}{.02in}}
\vspace{2cm}

In order to guage the success of the project, it is important to look at it's minimum requirements. The first requirement, a \underline{literature review} of the background material relevant to the project, was successfully met, and now makes up the 'Background Research' section of this report. The literature review proved extremely helpful in learning around the domain and even planning the approach to evaluating OpenStack. At least \underline{1 programmatic experiment plus 3 designed experiments} was also expected. This objective has been met and exceeded by far, with at least 6 re-usable experiments designed, with 5 of these being 'programmatic' i.e. using the REST APIs. Next, a \underline{Qualitative Evaluation report} was required, and this is delivered in part 1 of the 'Design \& Implementation' section, as a report on OpenStack architecture, design and capability; again, the target was successfully met. Similarly, a large \underline{re-usable software component} has been delivered in the form of a Spring based REST Client for OpenStack. \underline{Evaluation} of the project itself can be found in Chapter 4. \\
Clearly, the minimum requirements of this project, deemed reasonable by the project assessor, have been far exceeded in this project; furthermore, the production of an \underline{Experiment Framework}, listed as a possible extension, is built in to the REST client, meaning this extra objective was also achieved. The design of the Spring application which provides this can also be easily \underline{used with other configurations}, meaning that this project easily allows another possible extension, the running of experiments with different configurations; all that prevented this happening during the project was a lack of physical resources, and time to reconfigure the OpenStack environment. In terms of the Aims \& Objectives themselves, what is produced is very much in line with them, and the projected deliverables have all been matched and will be delivered with this report. It is for this reason that I would personally deem the project successful, based on it's initial, and in my opinion, ambitious aims. 